---
title: "Normalization and EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,error = FALSE,message = FALSE,fig.pos="h")
library(tidyverse)
library(limma)
library(ggrepel)
library(ggpubr)
source("R/utils.R")
```

# Exploration of quantitative data

The idea here is to answer the questions

- Is there significant variation in overall intensity between datasets and should this be normalized

```{r}
mq_data <- load_mq_annotated()
sample_data <- readxl::read_excel("raw_data/maxquant/sample_data.xlsx")
```

Boxplots of iBAQ and LFQ intensities show that there are quite large differences in intensity and that these are correlated with certain sample types.  

```{r}
mq_data_ibaq_matrix <- mq_data %>% 
  select(contains('iBAQ ')) %>% 
  log2 %>% 
  cbind(mq_data %>% select(prot_id)) %>% 
  gather(sample,intensity,-prot_id) %>% 
  add_column(measurement='iBAQ')


mq_data_lfq_matrix <- mq_data %>% 
  select(contains('LFQ intensity ')) %>% 
  log2 %>% 
  cbind(mq_data %>% select(prot_id)) %>% 
  gather(sample,intensity,-prot_id) %>% 
  add_column(measurement='LFQ')

mq_data_raw_intensities <- rbind(mq_data_ibaq_matrix,mq_data_lfq_matrix) %>% 
  mutate(sample = extract_sample_codes(sample)) %>% 
  filter(!is.na(intensity)) %>% 
  left_join(sample_data,by=c("sample"="sample_code"))

gp_rawi <- ggplot(mq_data_raw_intensities,aes(x=reorder(sample,sample_order),y=intensity)) + 
  geom_boxplot(aes(color=sample_type)) + 
  facet_wrap(~measurement) + 
  theme_pubclean() +
  theme(axis.text.x = element_text(size=8,angle=90), legend.title = element_blank(), legend.position = "bottom") + 
  xlab("") + ylab("Value")

ggsave(gp_rawi,filename = "figures/raw_boxplot.pdf",width=170,height=120, units = "mm")
```

LFQ measurements are probably suitable for normalization as they are already fairly close and since they are typically used for inter-sample comparisons. We don't normalize iBAQ since it is primarily for between protein comparisons.

Now we wish to make a PCA based on the normalized LFQ data.

```{r}

mq_data_lfq_lognorm_matrix <- lognorm_lfq(mq_data) %>% 
  select(prot_id,contains('LFQ intensity ')) %>% 
  column_to_rownames("prot_id") %>% 
  as.matrix()
  
mds <- plotMDS(mq_data_lfq_lognorm_matrix,plot = FALSE)



mds_data <- data.frame(x=mds$x,y=mds$y) %>% 
  rownames_to_column("sample") %>% 
  mutate(sample = extract_sample_codes(sample)) %>% 
  left_join(sample_data,by=c("sample"="sample_code"))

ggplot(mds_data,aes(x=x,y=y)) + 
  geom_point(aes(color=sample_type)) + 
  geom_label_repel(aes(label=sample))

p_pca <- ggplot(mds_data,aes(x=x,y=y)) + 
  geom_point(aes(color=sample_type)) + 
#  geom_label_repel(aes(label=sample),size=1) +
  theme_pubclean() +
  theme(legend.position = "right") +
  theme(text = element_text(size=6), legend.title = element_blank()) +
  theme(legend.text = element_text(size=6)) +
  xlab("Leading log2 fold change dim 1") +
  ylab("Leading log2 fold change dim 2") 
  
  

ggsave(p_pca,filename = "figures/sample_pca.pdf",width=80,height=60, units = "mm")

```

